---
title: "Homework 3"
author: JingYao Geng
date: '`r format(Sys.time(), "%Y-%m-%d") `'
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)

knitr::opts_chunk$set(
  fig.width = 6, 
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis", 
  ggplot2.continuous.fill = "viridis"
)

scale_cilour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

```{r Q1 setup}
data("instacart")
```

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns.

Observations are the level of items in orders by user. There are user / order variables -- user ID, order ID, order day, and order hour. There are also item variables -- name, aisle, department, and some numeric codes.

##### How many aisles are there, and which aisles are the most items ordered from?
```{r Q1_1}
instacart %>%
  count(aisle) %>% # 134 by 2, count all kind of aisles
  arrange(desc(n)) # arrange aisles by its amount, or reversely arrange(n)
```
* There are a total of 134 aisles in the instacart online grocery service with most items ordered from 'fresh vegetables' aisle.

##### Make a plot

Make a plot that shows the number of items ordered in each aisle, limiting this to aisles with more than 10000 items ordered. Arrange aisles sensibly, and organize your plot so others can read it.
```{r Q1_2}
instacart %>%
  count(aisle) %>% 
  filter(n > 10000) %>%
  mutate(
    aisle = factor(aisle), # a diff order: mutate to factor
    aisle = fct_reorder(aisle, n) # then, fct_reorder(var1, var2): reorder var1 by var2.
  ) %>%
  ggplot(aes(x = aisle, y = n)) +
  geom_point() +# google: ggplot rotate axis labels
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) # just this,looks not bad
```

##### Make a table

Make a table showing the three most popular items in each of the aisles “baking ingredients”, “dog food care”, and “packaged vegetables fruits”. Include the number of times each item is ordered in your table.
```{r Q1_3}
instacart %>%
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% # count product_name within the aisle
  mutate(rank = min_rank(desc(n))) %>% # min_rank() helps to rank the order
  filter(rank < 4) %>%
  arrange(aisle, rank) %>%
  # arrange(desc(n)) # rank as 123123123
  # arrange(rank): rank as 111222333
  knitr::kable() #printing table
```

##### Apples vs ice cream

Make a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week; format this table for human readers (i.e. produce a 2 x 7 table)
```{r Q1_4}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  pivot_wider(
    names_from = order_dow,
    values_from = mean_hour
  )
```
* On average, Coffee Ice Cream is ordered around 2:00 in the afternoon, and Pink Lady Apples is ordered about the noon.

### Problem 2

##### Load and tidy accelerometer data
```{r  Q2_import_and_clean, message=FALSE}
acc_df = 
  read_csv("./data/accel_data.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
    starts_with("activity_"),
    names_to = "activity_minutes", 
    names_prefix = "activity_",
    values_to = "activity_amounts") %>% 
  mutate(
    weekday_weekend = case_when(
      day %in% c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday") ~ "weekday",
      day %in% c("Sunday", "Saturday") ~ "weekend"),
    
    week = factor(week),
    day_id = as.factor(day_id),
    day = as.factor(day),
    activity_minutes = as.numeric(activity_minutes),
    activity_amounts = as.numeric(activity_amounts),
    weekday_weekend = as.factor(weekday_weekend),
    
    day = forcats::fct_relevel(day, c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
  ) %>%
  group_by(week) %>%
  arrange(day, .by_group = TRUE)
```
The original accelerometer dataset has 36 observations and 1443 columns. However, some of the variables are easy to understand. The tidy-version of the accelerometer dataset has `r nrow(acc_df)` rows and `r ncol(acc_df)` columns.

##### Tranditional analyses: create a table for totals
```{r Q2_total_var_table, message = FALSE}
acc_df %>%
  group_by(week,day) %>%
  summarize(total_activity = sum(activity_amounts)) %>%
  pivot_wider(
    names_from = day,
    values_from = total_activity) %>%
  knitr::kable(align = 'c')
```

The activity amounts were pretty stable for Wednesday and Thursday during the 5-week period. Saturday seems to be the day that the man had least activity amounts, especially for the last 2 weeks. Overall, the men has less activity amounts in week 4 and 5. It seems to be reasonable that it's hard to stick to one thing for a long time. Other trends might be difficult to find throughout the table.


##### Create a plot showing 24-hour activity time courses for each day
```{r Q2_plot}
acc_df %>% 
  group_by(day_id, day, week) %>%
  ggplot(aes(x = activity_minutes, y = activity_amounts, color = day, group = day_id)) + 
  stat_smooth(se = F, method = "loess") +

  labs(title = "24-hour Activity Inspection ",
       caption = "Data from accelerometer",
       x = "Hour of a Day",
       y = "Activity amounts") + 
  theme(plot.title = element_text(hjust = 0.5, size = 18)) +
  scale_x_continuous(
    breaks = seq(60,1440,60), 
    labels = as.character(c(1:24))) +
  scale_y_continuous(trans = "log")
  viridis::scale_color_viridis(discrete = TRUE)
```

We can see that most activities were during 8am to 9pm a day for the 5 weeks. The activity amounts were similar in most of the days, and the peaks of the activity amount usually happen around noon or in the afternoon. Other trends are hard to tell.


### Problem 3
```{r ny_noaa_start}
data("ny_noaa")
```
**A short description of the dataset**

The ny_noaa dataset contains `r ncol(ny_noaa)` variables and `r nrow(ny_noaa)` observation value. It displays the New York weather information from Jan. 1981 to Dec. 31 2010.

Important variables: 

* prcp: The daily precipitation volume (mm). 
* snow: The daily snowfall volume (mm).
* snwd: The daily snow depth (mm).
* tmax: Maximum temperature (degree C).
* tmin: Minimum temperature (degree C).

There are a lot of missing values in variable 'prcp', 'snow', 'snwd', 'tmax', and 'tmin'. In order to analyze the ny_noaa dataset and do some calculations based on the those variables, we need to remove all missing values.


# Read and clean the data,  
```{r 3.1_clean_and_tidy}
ny_df = ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(prcp = as.numeric(prcp, na.rm = TRUE),
         tmax = as.numeric(tmax, na.rm = TRUE),
         tmin = as.numeric(tmin, na.rm = TRUE),
         # reasonable units:
         prcp = prcp / 10,
         tmin = tmin / 10,
         tmax = tmax / 10)
```

* After the step of data clean and tidy, we organized the 'date' variable into 'year', 'month', and 'day' separately. Besides that, we convert variables 'prcp', 'tmin', and 'tmax' into mm, and degree C.

**For snowfall, what are the most commonly observed value?**
```{r 3.1_common_snowfall}
ny_df %>% 
  group_by(snow) %>%
  summarize(n = n()) %>% # count the snowfall volume
  filter(min_rank(desc(n)) < 2) # rank the count from large to small + the largest one
```

* We find the most commonly observed value in snowfall is 0mm. It's reasonable that snow only happens during winter session.

**Average temperature in January and July (across station) in year 1981 and 2010.** 
```{r 3.2_mean}
temp_mean <- 
  ny_df %>%
  select(id, year, month, day, tmax, tmin) %>% 
  filter( month ==  "01" | month == "07" ) %>% 
  group_by(id, year, month) %>% 
  summarise(
    mean_tmax = mean(tmax, na.rm = T)) %>% 
    drop_na(mean_tmax)
```

**The max temp plots** 
```{r 3.2_max_temp_plots}
ggplot(temp_mean, aes(x = as.numeric(year), y = as.numeric(mean_tmax), color = id )) +
  geom_point(alpha = 0.5) +
  scale_x_continuous(breaks = c(1981, 1985, 1990, 1995, 2000, 2005, 2010),
                  labels = c("1981", "1985", "1990", "1995", "2000", "2005", "2010")) +
  scale_y_continuous(breaks = c(-10, 0, 10, 20, 30),
                     labels = c("-10°C", "0°C", "10°C", "20°C", "30°C")) +
  labs(
    title = "Max tempreture in NY weather stations in January and July (1991-2010)",
    x = "Year",
    y = "Average Monthly Tempreture (°C)") + 
  theme(legend.position = "none")  +
  facet_grid(~month) +
  viridis::scale_color_viridis(discrete = TRUE, option = "plasma") 
```

* By comparing the two plots, we find that the average temperature in January is always lower than the average temperature in July. It's true that the former is in the winter session and the latter is in the summer session.

* The January plot might somehow indicate the existence of global warming as the gradually increased average temperature from 1981 to 2010 in this plot. 

* Several outliers: 
  * There is one station that was extremely cool in 1982 in plot "01".
  * There is one station that was not warm as usual in 1988 in plot "07".




